---
title: "Homework 2"
subtitle: "Research Methods, Spring 2025"
author: Justin Hahm
format:
  pdf:
    output-file: "hahm-j-hwk2-1"
    output-ext: "pdf"
    header-includes:
      - \usepackage{float}
      - \floatplacement{table}{H}
---

```{r}
#| include: false

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes, fixest,
               scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD,
               ggrepel, survey, foreign, devtools, pdftools, kableExtra, modelsummary,
               kableExtra)
```



```{r}
#| include: false
#| eval: true
 
load("C:/Users/hjust/OneDrive/Documents/GitHub/homework2/submission1/Hwk2_workspace.Rdata")
```


My answers to the homework questions are described below. Note that I do the analysis for these answers in a separate `R` script. My analysis file is available in the analysis folder. The GitHub repository for this work is available [here](https://github.com/HJHAHM1/homework2.git). Enjoy!


\newpage
# Summarize the Data
\noindent 1. How many hospitals filed more than one report in the same year? Show your answer as a line graph of the number of hospitals over time. <br>

```{r} 
#| echo: false

ggplot(full.hcris.data.multiple, aes(x = fyear, y = duplicates)) +
  geom_line() +  
  geom_point() +  
  labs(title = "Number of Hospitals With More than One Report per Year Over Time",
       x = "Fiscal Year",
       y = "Number of Hospitals") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))
```

\newpage
\noindent 2. After removing/combining multiple reports, how many unique hospital IDs (Medicare provider numbers) exist in the data?
```{r}
#| echo: false

full.hcris.data %>% group_by(year) %>%
  summarize(hosp_count = n()) %>%
  ggplot(aes(x=as.factor(year),y=hosp_count, group = 1))+
  geom_line()+
  labs(
    x= "Year",
    y="Number of Hospitals")+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))

```

\newpage
\noindent 3. What is the distribution of total charges (tot_charges in the data) in each year? Show your results with a “violin” plot, with charges on the y-axis and years on the x-axis.
```{r}
#| echo: false
ggplot(final_data_clean, aes(x = factor(year), y = log_charges)) +
  geom_violin(trim = FALSE, fill = "lightblue", color = "black") 
  labs(title = "Distribution of Total Charges by Year",
       x = "Year",
       y = "log(Total Charges)") +
  theme_minimal() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8), 
    axis.text.y = element_text(size = 8), 
    axis.title.x = element_text(size = 10), 
    axis.title.y = element_text(size = 10),  
    plot.title = element_text(size = 12))

```

\newpage
\noindent 4. What is the distribution of estimated prices in each year? Again present your results with a violin plot, and recall our formula for estimating prices from class. Be sure to do something about outliers and/or negative prices in the data.
```{r}
#| echo: false
ggplot(full.hcris.data_filtered, aes(x = factor(year), y = price)) +
  geom_violin(fill = "lightblue", color = "darkblue") +
  theme_minimal() +
  labs(title = "Distribution of Estimated Prices by Year",
       x = "Year",
       y = "Price") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


\newpage
# Estimate ATEs

\noindent 5. Calculate the average price among penalized versus non-penalized hospitals.

```{r}
#| echo: false
#| label: tbl-pen
#| tbl-cap: Average price among penalized vs non-penalized hospitals
options(knitr.kable.NA = 0)
knitr::kable(mean_values, digits=0)
```

\newpage
\noindent 6. Split hospitals into quartiles based on bed size. To do this, create 4 new indicator variables, where each variable is set to 1 if the hospital’s bed size falls into the relevant quartile. Provide a table of the average price among treated/control groups for each quartile.
```{r}
#| echo: false
#| label: tbl-quart
#| tbl-cap: Quartiles by bed size in 2012

options(knitr.kable.NA = 0)
knitr::kable(summary_table, digits=0,
             col.names=c("Quartile","Nonpenalty","Penalty"),
             format.args=list(big.mark=","), booktabs = TRUE)
```

\newpage
\noindent 7. Find the average treatment effect using each of the following estimators, and present your results in a single table:
```{r}
#| echo: false
#| label: tbl-ate
#| tbl-cap: ATE 

options(knitr.kable.NA = 0)
knitr::kable(ate_summary, digits=2,
             col.names=c("Method","ATE"),
             format.args=list(big.mark=","), booktabs = TRUE)
```

\newpage
\noindent 8. With these different treatment effect estimators, are the results similar, identical, very different?

\indent The results from the different treatment effect estimators are identical. 

\noindent 9. Do you think you’ve estimated a causal effect of the penalty? Why or why not? (just a couple of sentences)

\indent I believe I have estimated a causal effect of the penalty becuase I see convergence across different methods. This indicates consistency. The selection on observables assumption should be satisfied. 

\noindent 10. Briefly describe your experience working with these data (just a few sentences). Tell me one thing you learned and one thing that really aggravated or surprised you.

\indent It was difficult to keep track of the different variable names and the dfs after filtering. While working with this dataset, I realized the importance of being organized and working through my code sequentially. The values would change if I ran the code again out of order while debugging, which became frustrating.